---
title: "Automated Maschine Learning with H2O"
author: "Christian Sühl"
---

```{r}
library(h2o)

# To launch H2O locally with default initialization arguments, use the following: 
h2o.init()
```

```{r}
library(tidyverse)
car_stats <- function(groupby_var, measure_var) {
    groupby_var <- enquo(groupby_var)
    measure_var <- enquo(measure_var)
    ret <- mtcars %>% 
             group_by(!!groupby_var) %>%
             summarize(min = min(!!measure_var), max = max(!!measure_var)) %>%
             # Optional: as_label() and "walrus operator" :=
             mutate(measure_var = as_label(measure_var), !!measure_var := "test")
    return(ret)
}
car_stats(cyl,hp)
scatter_plot <- function(data, x_var, y_var) {
  x_var <- enquo(x_var)
  y_var <- enquo(y_var)
  ret <- data %>% 
           ggplot(aes(x = !!x_var, y = !!y_var)) + 
           geom_point() + 
           geom_smooth() +
           ggtitle(str_c(as_label(y_var), " vs. ",as_label(x_var)))
  return(ret)
}
scatter_plot(mtcars, cyl, hp)
```

```{r}
library(readr)
# Load data
employee_attrition_tbl <- read_csv("datasets-1067-1925-WA_Fn-UseC_-HR-Employee-Attrition.csv")

# Business & Data Understanding: Department and Job Role

# Data subset
dept_job_role_tbl <- employee_attrition_tbl %>%
  select(EmployeeNumber, Department, JobRole, PerformanceRating, Attrition)

dept_job_role_tbl %>%

  group_by(Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

# Attrition by department
dept_job_role_tbl %>%

  # Block 1
  group_by(Department, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%

  # Block 2: Caution: It's easy to inadvertently miss grouping when creating counts & percents within groups
  group_by(Department) %>%
  mutate(pct = n / sum(n))

# Attrition by job role
dept_job_role_tbl %>%

  # Block 1
  group_by(Department, JobRole, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%

  # Block 2
  group_by(Department, JobRole) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%

  # Block 3
  filter(Attrition %in% "Yes")
```

```{r}
# Use this
# Function to convert counts to percentages. 
count_to_pct <- function(data, ..., col = n) {
  # capture the dots
  grouping_vars_expr <- quos(...)
  col_expr <- enquo(col)
  ret <- data %>%
    group_by(!!! grouping_vars_expr) %>%
    mutate(pct = (!! col_expr) / sum(!! col_expr)) %>%
    ungroup()
  return(ret)
}

assess_attrition <- function(data, attrition_col, attrition_value, baseline_pct) {

  attrition_col_expr <- enquo(attrition_col)

  data %>%
  
    # Use parenthesis () to give tidy eval evaluation priority
    filter((!! attrition_col_expr) %in% attrition_value) %>%
    arrange(desc(pct)) %>%
    mutate(
      # Function inputs in numeric format (e.g. baseline_pct = 0.088 don't require tidy eval)
      above_industry_avg = case_when(
        pct > baseline_pct ~ "Yes",
        TRUE ~ "No"
      )
    )

}

# Function to calculate attrition cost
calculate_attrition_cost <- function(

  # Employee
  n                    = 1,
  salary               = 80000,

  # Direct Costs
  separation_cost      = 500,
  vacancy_cost         = 10000,
  acquisition_cost     = 4900,
  placement_cost       = 3500,

  # Productivity Costs
  net_revenue_per_employee = 250000,
  workdays_per_year        = 240,
  workdays_position_open   = 40,
  workdays_onboarding      = 60,
  onboarding_efficiency    = 0.50

) {

  # Direct Costs
  direct_cost <- sum(separation_cost, vacancy_cost, acquisition_cost, placement_cost)

  # Lost Productivity Costs
  productivity_cost <- net_revenue_per_employee / workdays_per_year *
    (workdays_position_open + workdays_onboarding * onboarding_efficiency)

  # Savings of Salary & Benefits (Cost Reduction)
  salary_benefit_reduction <- salary / workdays_per_year * workdays_position_open

  # Estimated Turnover Per Employee
  cost_per_employee <- direct_cost + productivity_cost - salary_benefit_reduction

  # Total Cost of Employee Turnover
  total_cost <- n * cost_per_employee

  return(total_cost)

}

# Function to plot attrition
plot_attrition <- function(data, 
                           ..., 
                           .value,
                           fct_reorder = TRUE,
                           fct_rev     = FALSE,
                           include_lbl = TRUE,
                           color       = "#2dc6d6",
                           units       = c("0", "K", "M")) {

  ### Inputs
  group_vars_expr   <- quos(...)
  
  # If the user does not supply anything, 
  # this takes the first column of the supplied data
  if (length(group_vars_expr) == 0) {
    group_vars_expr <- quos(rlang::sym(colnames(data)[[1]]))
    }

  value_expr <- enquo(.value)

  units_val  <- switch(units[[1]],
                       "M" = 1e6,
                       "K" = 1e3,
                       "0" = 1)
  if (units[[1]] == "0") units <- ""

  # Data Manipulation
  # This is a so called Function Factory (a function that produces a function)
  usd <- scales::dollar_format(prefix = "$", largest_with_cents = 1e3)

  # Create the axis labels and values for the plot
  data_manipulated <- data %>%
    mutate(name = str_c(!!! group_vars_expr, sep = ": ") %>% as_factor()) %>%
    mutate(value_text = str_c(usd(!! value_expr / units_val),
                              units[[1]], sep = ""))

  
  # Order the labels on the y-axis according to the input
  if (fct_reorder) {
    data_manipulated <- data_manipulated %>%
      mutate(name = forcats::fct_reorder(name, !! value_expr)) %>%
      arrange(name)
  }

  if (fct_rev) {
    data_manipulated <- data_manipulated %>%
      mutate(name = forcats::fct_rev(name)) %>%
      arrange(name)
  }

  # Visualization
  g <- data_manipulated %>%

        # "name" is a column name generated by our function internally as part of the data manipulation task
        ggplot(aes(x = (!! value_expr), y = name)) +
        geom_segment(aes(xend = 0, yend = name), color = color) +
        geom_point(aes(size = !! value_expr), color = color) +
        scale_x_continuous(labels = scales::dollar) +
        scale_size(range = c(3, 5)) +
        theme(legend.position = "none")

  # Plot labels if TRUE
  if (include_lbl) {
    g <- g +
      geom_label(aes(label = value_text, size = !! value_expr),
                 hjust = "inward", color = color)
  }

  return(g)

}

dept_job_role_tbl %>%

  # Block 1
  count(Department, JobRole, Attrition) %>%

  # Block 2
  count_to_pct(Department, JobRole) %>%  

  # Block 3
  assess_attrition(Attrition, attrition_value = "Yes", baseline_pct = 0.088) %>%

  # Block 4. Set salaray to 80000 for now
  mutate(
    cost_of_attrition = calculate_attrition_cost(n = n, salary = 80000)
  ) %>%

  # Select columnns
  plot_attrition(Department, JobRole, .value = cost_of_attrition,
                 units = "M") +
  labs(
    title = "Estimated Cost of Attrition by Job Role",
    x = "Cost of Attrition",
    subtitle = "Looks like Sales Executive and Labaratory Technician are the biggest drivers of cost"
  )
```

```{r}
# Libraries 
library(tidyverse)
library(readxl)
library(skimr)
library(GGally)

# Load Data data definitions

path_data_definitions <- "data_definitions.xlsx"
definitions_raw_tbl   <- read_excel(path_data_definitions, sheet = 1, col_names = FALSE)

employee_attrition_tbl

# Descriptive Features
employee_attrition_tbl %>% select(Age, DistanceFromHome, Gender, MaritalStatus, NumCompaniesWorked, Over18)

# Employment Features
employee_attrition_tbl %>% select(Department, EmployeeCount, EmployeeNumber, JobInvolvement, JobLevel, JobRole, JobSatisfaction)

# Compensation Features
employee_attrition_tbl %>% select(DailyRate, HourlyRate, MonthlyIncome, MonthlyRate, PercentSalaryHike, StockOptionLevel)

# Survery Results
employee_attrition_tbl %>% select(EnvironmentSatisfaction, JobSatisfaction, RelationshipSatisfaction, WorkLifeBalance)

# Performance Data
employee_attrition_tbl %>% select(JobInvolvement, PerformanceRating)

# Work-Life Features
employee_attrition_tbl %>% select(BusinessTravel, OverTime)

# Training & Education
employee_attrition_tbl %>% select(Education, EducationField, TrainingTimesLastYear)

# Time-Based Features
employee_attrition_tbl %>% select(TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)
```

```{r}
# Step 1: Data Summarization -----

skim(employee_attrition_tbl)

# Character Data Type
employee_attrition_tbl %>%
    select_if(is.character) %>%
    glimpse()

# Get "levels"
employee_attrition_tbl %>%
    select_if(is.character) %>%
    map(unique)

# Proportions    
employee_attrition_tbl %>%
    select_if(is.character) %>%
    map(~ table(.) %>% prop.table())

# Numeric Data
employee_attrition_tbl %>%
    select_if(is.numeric) %>%
    map(~ unique(.) %>% length())

employee_attrition_tbl %>%
    select_if(is.numeric) %>%
    map_df(~ unique(.) %>% length()) %>%
    # Select all columns
    pivot_longer(everything()) %>%
    arrange(value) %>%
    filter(value <= 10)
```

```{r}
library(GGally)
# Step 2: Data Visualization ----

# Create data tibble, to potentially debug the plot_ggpairs function (because it has a data argument)
data <- employee_attrition_tbl %>%
    select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome)

plot_ggpairs <- function(data, color = NULL, density_alpha = 0.5) {
    
    color_expr <- enquo(color)
    
    if (rlang::quo_is_null(color_expr)) {
        
        g <- data %>%
            ggpairs(lower = "blank") 
        
    } else {
        
        color_name <- quo_name(color_expr)
        
        g <- data %>%
            ggpairs(mapping = aes_string(color = color_name), 
                    lower = "blank", legend = 1,
                    diag = list(continuous = wrap("densityDiag", 
                                                  alpha = density_alpha))) +
            theme(legend.position = "bottom")
    }
    
    return(g)
    
}

employee_attrition_tbl %>%
    select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
    plot_ggpairs(color = Attrition)
```

```{r}
# Explore Features by Category

#   1. Descriptive features: age, gender, marital status 
employee_attrition_tbl %>%
    select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
    plot_ggpairs(Attrition)

#   2. Employment features: department, job role, job level
employee_attrition_tbl %>%
    select(Attrition, contains("employee"), contains("department"), contains("job")) %>%
    plot_ggpairs(Attrition) 

#   3. Compensation features: HourlyRate, MonthlyIncome, StockOptionLevel 
employee_attrition_tbl %>%
    select(Attrition, contains("income"), contains("rate"), contains("salary"), contains("stock")) %>%
    plot_ggpairs(Attrition)

#   4. Survey Results: Satisfaction level, WorkLifeBalance 
employee_attrition_tbl %>%
    select(Attrition, contains("satisfaction"), contains("life")) %>%
    plot_ggpairs(Attrition)

#   5. Performance Data: Job Involvment, Performance Rating
employee_attrition_tbl %>%
    select(Attrition, contains("performance"), contains("involvement")) %>%
    plot_ggpairs(Attrition)

#   6. Work-Life Features 
employee_attrition_tbl %>%
    select(Attrition, contains("overtime"), contains("travel")) %>%
    plot_ggpairs(Attrition)

#   7. Training and Education 
employee_attrition_tbl %>%
    select(Attrition, contains("training"), contains("education")) %>%
    plot_ggpairs(Attrition)

#   8. Time-Based Features: Years at company, years in current role
employee_attrition_tbl %>%
    select(Attrition, contains("years")) %>%
    plot_ggpairs(Attrition)
```


# Challenge
## Question 1 | What can you deduce about the interaction between Monthly Income and Attrition?
There is a big spike of people having attrition at the low end of the monthly income spectrum far greater than people having no attrition with low monthly income. This indicates that monthly income has a relatively strong affect on attrition.
## Question 2 | What can you deduce about the interaction between Percent Salary Hike and Attrition?
The attrition graph depending on the percent salary hike is nearly identical for people with attrition and without. Meaning a percent salary hike has very little effect on attrition.
## Question 3 | What can you deduce about the interaction between Stock Option Level and Attrition?
The attrition vs. stock option level graph shows that people having a stock option level of 1 or to a lesser extend 2 are much less likely to have high attrition, meaning there could be a positive effect between the stock option level and attrition for the employer.
## Question 4 | What can you deduce about the interaction between Environment Satisfaction and Attrition?
First of all, the satisfaction levels of the survey seemed to be clumbed in certain levels which is likely the result of some kind of numerical scale being used for this survey (Rate from 1 to 4 for example). Considering that, there is an upward trend in the number of people having no attrition with higher satisfaction.
## Question 5 | What can you deduce about the interaction between Work Life Balance and Attrition
The work life balance seems to indicate a stronger connection with attrition, with people rating their work life balance more positively having less attrition.
## Question 6 | What Can you deduce about the interaction between Job Involvement and Attrition?
People that are more involved in the job are less likely to have attrition.
## Question 7 | What can you deduce about the interaction between Over Time and Attrition?
Roughly the same amount of poeple that have attrition do and do not perform overtime. But as a percentage, the poeple doing overtime are much more likely to have attrition.
## Question 8 | What can you deduce about the interaction between Training Times Last Year and Attrition
There is a slight trend, that more training time results in less attrition but as far as I can tell very minor.
## Question 9 | What can you deduce about the interaction between Years At Company and Attrition
Regarding years at the company for very new people the attrition is much higher than for those that are very long with the company.
Crucially those that stay long have very low attrition, since most likely those with high attrition left at some point, so the data is inherently somewhat skewed.
## Question 10 | What can you deduce about the interaction between Years Since Last Promotion and Attrition?
Years since last promotion seem to have a rather small affect on attrition but the small bump in attrition if people are longer with the company is notworthy, meaning that people that stayed with the company for an extended amount of time without a promotion start to get more attrition.
